Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2021-08-15 numerics_paper/version2/numerical_chj.tex

{\color{red}
  Finally, following \cite[Section III B]{kitamura2009evaluation}, we repeat the same experiment using
  the blended solver with an initial condition where the shock,
  located at $r_s=0.3$, is aligned with a grid interface. Afterward, we perturb the initial condition near the shock to misalign it
  from the grid interface. We use a mesh with $90$ elements in the radial direction and $90$ elements in the angular direction;
  see Figure \ref{fig:mesh_chj}.
  For this mesh, $\Delta r=(r_\out-r_\jet)/90=0.01$. 
  Let $(x_i,y_i)$ denote the coordinates of the center of the $i$-th element and $r_i:=\sqrt{x_i^2+y_i^2}$.
  For all elements where $r_i\not\in[r_s,r_s+\Delta r]$, we set
  \begin{align*}
    h_i = \frac{1}{\Delta r}\int_{r_i-\Delta r/2}^{r_i+\Delta r/2}h(r) dr,
    \qquad
    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
    \qquad
    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
  \end{align*}
  where $\beta=r_\jet h_\jet |u_\jet|$ and $h(r)$ is the solution of the ODE \eqref{steady} from either
  $r=r_\jet$ to $r=r_s^-$ or $r_s^+$ to $r_\out$.
  If $r_i = r_s+\Delta r/2$, we set
  \begin{align*}
    h_i = \delta h_L + (1+\delta) h_R,
    \qquad
    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
    \qquad
    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
  \end{align*}
  where $\delta=\{0,0.1,0.2,\dots,0.9,1\}$ and $h_L$ and $h_R$ are the values of the water height
  at the left and right cells from the cell with $r_i=r_s+\Delta r/2$, respectively.
  In Figure \ref{stab_shock_init_cond}, we plot $h(r,\theta_j)$ with $j=\{1,2,\dots,90\}$; that is,
  we plot the water height as a function of $r$ for every value of $\theta$.
  In Figures \ref{stab_shock_roe} and \ref{stab_shock_blended}, we plot the corresponding solutions at
  $t=3$ using method \eqref{second-order_via_fluct} with the Roe solver and with the blended solver,
  respectively. With the Roe solver, the carbuncle instability appears; as a consequence, the
  location of the shock changes for the different values of $\theta$.
  With the blended solver, no carbuncle instability appears and, therefore, all slices are aligned.
  In the figures we only show the results using $\delta=\{0,0.5,1\}$ but we
  obtained similar results with the other values of $\delta$.

  \begin{figure}
    \subfloat[Initial condition.\label{stab_shock_init_cond}]{
    \begin{tabular}{c}
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0.png} 
    \end{tabular}
    }
    ~
    \subfloat[Roe's solver.\label{stab_shock_roe}]{
    \begin{tabular}{c}
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_roe.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_roe.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_roe.png} 
    \end{tabular}
    }
    ~
    \subfloat[Blended solver.\label{stab_shock_blended}]{
    \begin{tabular}{c}
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_blended.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_blended.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_blended.png} 
    \end{tabular}
  }    
    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
      We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
      In (a) we show the different initial conditions that we consider.
      In (b) and (c) we show the corresponding solutions at $t=3$ using method
      \eqref{second-order_via_fluct} with the Roe solver and with the blended solver, respectively. 
      \label{fig:stab_shock}}
  \end{figure}
}

>>>>

{\color{red}
  Next we perform a set of experiments similar to those of \cite[Section III B]{kitamura2009evaluation}.
  In these experiments, we perturb the initial condition in just one layer of cells, immediately behind the shock.
  We use an annular mesh (see Figure \ref{fig:mesh_chj}) with $90$ elements in
  the radial direction and $90$ elements in the angular direction.  Let $h(r)$ denote the exact solution
  of the ODE \eqref{steady}, and let $r_i$ denote the radius of a given cell center.  Then for all cells
  away from the shock (i.e. those with $r_i \notin [0.3+\Delta r]$) we use the exact solution averaged over the grid:
  \begin{align*}
    h_i = \frac{1}{\Delta r}\int_{r_i-\Delta r/2}^{r_i+\Delta r/2}h(r) dr,
    \qquad
    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
    \qquad
    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
  \end{align*}
  where $\beta=r_\jet h_\jet |u_\jet|$ and $(x_i,y_i)$ are the cartesian
  coordinates of the cell center.  For the layer of cells just behind the shock (with $0.3 \le r \le 0.3+\Delta r$),
  we set
  \begin{align*}
    h_i = \delta h_{i-1} + (1+\delta) h_{i+1},
    \qquad
    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
    \qquad
    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
  \end{align*}
  where $\delta=\{0,0.1,0.2,\dots,0.9,1\}$ and $h_L$ and $h_R$ are the values of the water height
  at the left and right cells from the cell with $r_i=r_s+\Delta r/2$, respectively.
  In Figure \ref{stab_shock_init_cond}, we plot representative results of these experiments, showing
  the initial data in the first column and solutions at $t=3$ in the second and third columns.
  we show only the results using $\delta=\{0,0.5,1\}$ since other results were similar.
  The solution obtained with the Roe solver, in which the carbuncle instability appears,
  is sensitive to the perturbation just described, while the solution obtained with the
  blended solver is radially symmetric and essentially the same for all values of $\delta$.
  \begin{figure}
    \subfloat[Initial condition ($t=0$)\label{stab_shock_init_cond}]{
    \begin{tabular}{c}
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0.png} 
    \end{tabular}
    }
    ~
    \subfloat[Roe's solver ($t=3$)\label{stab_shock_roe}]{
    \begin{tabular}{c}
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_roe.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_roe.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_roe.png} 
    \end{tabular}
    }
    ~
    \subfloat[Blended solver ($t=3$)\label{stab_shock_blended}]{
    \begin{tabular}{c}
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_blended.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_blended.png} \\ \\
      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_blended.png} 
    \end{tabular}
  }    
    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1},
      using the 2nd-order method \eqref{second-order_via_fluct} with two Riemann solvers. The depth $h$
      is shown as a function of $r$.  Note that the plots in the third column contain include lines for
      each angle along the grid, but only a single line is visible due to the solution's symmetry.
      \label{fig:stab_shock}}
  \end{figure}
}
